---
title: "Day 5 Homework"
output: html_document
---

0. Reload your data, and re-use the code you already have for getting things formatted nicely (columns renamed, factors relabeled, etc.)

```{r}
library(dplyr)
prof.salary <- read.csv("salaries.csv")
head(prof.salary)
colnames(prof.salary)[1] <- "ID"
prof.salary$ID <- as.character(prof.salary$ID)
summary(prof.salary)

```

1. Aggregate (e.g., get mean/median/max/something) of some numerical variable, over one or more factors.  Use `group_by()` plus `summarize()` from the `dplyr` package. You may or may not need to reshape the data first!

```{r}
salary.means <- summarize(group_by(prof.salary, rank, sex), mean.salary = mean(salary), 
                          std.salary = sd(salary))
salary.means
```

2. Get your data into "long" (aka "tidy") format, if it isn't already

```{r}
# I want to practice gather() in the following code, but I am not sure whether it is necessary in real practice. 
salary.long <- gather(prof.salary, time.period, yrs, 4:5)
salary.long$time.period <- as.factor(salary.long$time.period)
levels(salary.long$time.period) <- c("service", "since phd")
head(salary.long)
```







